### 2023-08-31 reports/paper/paper.qmd
remove names

**Authors:** Kevin T. Chen\^a, b\^, Sally Picciotto^c^, Ellen A.
Eisen^c^

@@

**Running head suggestion:** Hypothetical interventions with guaranteed positivity

@@

**Word count** **(Limit: 1500 to 4000 words):** 3842 (approx)

@@

(Limit: 250 words)

**Background:** Non-Hodgkin lymphoma (NHL) incidence has increased
substantially in the US and was recently linked to workplace exposure to
soluble metalworking fluids (MWF) in a standard survival analysis of the
United Auto Workers-General Motors (UAW-GM) cohort. Here, we estimate
the effects of hypothetical interventions on soluble MWF exposure in
relation to NHL risk in the same cohort of autoworkers. Estimation of
causal effects requires positivity, which we guarantee by investigating
causal contrasts that are supportable by the data.

**Methods:** We defined hypothetical supportable interventions that
reduce annual daily average exposures above a specific target limit to
that limit when supported by data. When the target limit is not
supported, the intervention reduces exposure to the nearest supported
level of exposure. We estimated counterfactual risk of NHL between 1985
and 2015 in the UAW-GM cohort of autoworkers at three Michigan plants (n
= 33,134) under hypothetical supportable interventions on exposure to
soluble MWF.

**Results:** 339 NHL cases occurred over the 30-year follow-up period.
Stronger target limits on soluble MWF exposure resulted in monotonic
reductions in NHL risk. Setting the target exposure limit at 0.5
mg/m^3^, the NIOSH recommended exposure limit, would have prevented 124
(95% CI: 66, 202) cases.

**Conclusion:** More stringent limits on exposure to soluble MWF provide
stronger protections against NHL. We expect that uniformly enforcing the
target exposure limits, regardless of data support, would have even
stronger protective effects. Strengthening exposure limits for MWF may
protect workers from NHL during the anticipated boom in domestic
manufacturing.

>>>>

**Authors:** Kevin T. Chen^a,b^, Sally Picciotto^c^, Ellen A. Eisen^c^

@@

@@

**Word count** **(Limit: 4000 words):** 3842

@@

(Limit: 200 words)

Non-Hodgkin lymphoma (NHL) incidence was recently linked to exposure to
metalworking fluid (MWF) in a standard survival analysis of the United
Auto Workers-General Motors cohort. To provide results more directly
relevant to guiding exposure limits, we estimated the counterfactual
risks of NHL under hypothetical supportable interventions on exposure to
MWF in the same cohort (n = 33,134) for 1985-2015 using the
hazard-extended iterative conditional expectation g-formula. We
addressed potential bias due to the healthy worker survivor effect while
investigating stochastic dynamic interventions that avoid extrapolation
beyond observed conditional exposure distributions, thus guaranteeing
positivity by design. These interventions reduced exposures above
specified target limits to the nearest supported level of exposure but
allowed exposures below the limit to vary naturally. 339 NHL cases
occurred over the 30-year follow-up period. Stronger target limits on
MWF exposure resulted in monotonic reductions in NHL risk. Setting the
target exposure limit at 0.5 mg/m^3^, the NIOSH recommended exposure
limit, would have prevented 124 (95% CI: 66, 202) cases. We expect that
uniformly enforcing the target exposure limits, regardless of data
support, would have yielded even stronger protective effects.
Strengthening protections against exposure to MWF may protect workers
from NHL during the anticipated boom in domestic manufacturing.
